---
title: "Grouping OLS predictions"
output: html_notebook
---

Philip suggested I should try grouping the results I had using OLS regressions
into the 4 groups, and see if there is any difference there.

```{r}
library(gdata)
library(nnet)
load('~/data/baseline_prediction/combined_descriptives_12172018.RData.gz')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
df = merge(clin, data, by='MRN')
idx = df$diag_group != 'new_onset'
idx2 = !is.na(df$inatt_vol_lh) | !is.na(df$inatt_AD_clu1) | !is.na(df$inatt_melodic_DMN)
imaging = df[idx & idx2,]
imaging[imaging$OLS_inatt_slope <= -.33, 'OLS_inatt_categ'] = 'marked'
imaging[imaging$OLS_inatt_slope > -.33 & imaging$OLS_inatt_slope <= 0, 'OLS_inatt_categ'] = 'mild'
imaging[imaging$OLS_inatt_slope > 0, 'OLS_inatt_categ'] = 'deter'
imaging[imaging$DX == 'NV', 'OLS_inatt_categ'] = 'NV'
imaging$OLS_inatt_categ = as.factor(imaging$OLS_inatt_categ)
imaging$OLS_inatt_categ = relevel(imaging$OLS_inatt_categ, ref='NV')
load('~/data/baseline_prediction/combined_descriptives_12172018.RData.gz')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
df = merge(clin, data, by='MRN')
idx = df$diag_group != 'new_onset'
struct = df[!is.na(df$HI_vol_rh) & idx,]
load('~/data/baseline_prediction/struct_volume_11142018_260timeDiff12mo.RData.gz')
struct = merge(struct, data, by='MRN') # put mask ids in combined dataset
mprage = read.xls('~/data/baseline_prediction/long_scans_08072018.xlsx',
                  sheet='mprage')
struct = merge(struct, mprage, by.x='mask.id', by.y='Mask.ID...Scan') # get demographics
qc = read.csv('~/data/baseline_prediction/master_qc.csv')
struct = merge(struct, qc, by.x='mask.id', by.y='Mask.ID') # get QC scores
df = merge(struct, imaging, by='MRN')
dim(df)

pairwise.t.test(df$inatt_vol_lh.x, df$OLS_inatt_categ, p.adjust.method='none')

fm = as.formula(df$inatt_vol_lh.x ~ df$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('volume LH inatt, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))

```

So, the idea now is to regress out some of these significant variables to see what happens, and also try to make the NVs a bit more pure (i.e. flat OLS only).

```{r}
bad_nvs = df$OLS_inatt_categ == 'NV' & df$OLS_inatt_slope.x != 0
df_clean = df[!bad_nvs, ]
dim(df_clean)
```

Now we figure out how the cluster looks like without any possible significant variables:

```{r}
library(MASS)
res.lm <- lm(inatt_vol_lh.x ~ Sex...Subjects + int_avg_freesurfer5.3 + age_at_scan + I(age_at_scan^2) + ext_avg_freesurfer5.3 + int_avg_freesurfer5.3 + mprage_QC, data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_inatt_categ)))
pairwise.t.test(mycluster, df_clean$OLS_inatt_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('volume LH inatt, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

And of course the last thing is to run the logistic model:

```{r}
library(nnet)
fit <- multinom(df_clean$OLS_inatt_categ ~ scale(mycluster), na.action=na.omit)
z <- summary(fit)$coefficients/summary(fit)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
rr = exp(coef(fit))
pp = fitted(fit)
print(p)
print(fit)
print(rr)
```

The only significant category when compared to NVs is marked improvement:

* A one-unit increase in the volume of the left hemisphere cluster variable (i.e. increase by 1 SD) is associated with an increase in the log odds of markedly improving (vs normals) in the amount of .41.
* In terms of relative risk ratio, a one-unit increase in the the volume of that brain cluster yields a relative risk ratio of 1.5 of marked improvement. In other words, the odds of marked improvement, compared to normals, is 1.5 times higher for every 1 SD we increase in that brain region.

We should also look at the right hemisphere results as well. Those were for HI though:

```{r}
load('~/data/baseline_prediction/combined_descriptives_12172018.RData.gz')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
df = merge(clin, data, by='MRN')
idx = df$diag_group != 'new_onset'
idx2 = !is.na(df$inatt_vol_lh) | !is.na(df$inatt_AD_clu1) | !is.na(df$inatt_melodic_DMN)
imaging$OLS_HI_categ = NULL
imaging[imaging$OLS_HI_slope <= -.5, 'OLS_HI_categ'] = 'marked'
imaging[imaging$OLS_HI_slope > -.5 & imaging$OLS_HI_slope <= 0, 'OLS_HI_categ'] = 'mild'
imaging[imaging$OLS_HI_slope > 0, 'OLS_HI_categ'] = 'deter'
imaging[imaging$DX == 'NV', 'OLS_HI_categ'] = 'NV'
imaging$OLS_HI_categ = as.factor(imaging$OLS_HI_categ)
imaging$OLS_HI_categ = relevel(imaging$OLS_HI_categ, ref='NV')
load('~/data/baseline_prediction/combined_descriptives_12172018.RData.gz')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
df = merge(clin, data, by='MRN')
idx = df$diag_group != 'new_onset'
struct = df[!is.na(df$HI_vol_rh) & idx,]
load('~/data/baseline_prediction/struct_volume_11142018_260timeDiff12mo.RData.gz')
struct = merge(struct, data, by='MRN') # put mask ids in combined dataset
mprage = read.xls('~/data/baseline_prediction/long_scans_08072018.xlsx',
                  sheet='mprage')
struct = merge(struct, mprage, by.x='mask.id', by.y='Mask.ID...Scan') # get demographics
qc = read.csv('~/data/baseline_prediction/master_qc.csv')
struct = merge(struct, qc, by.x='mask.id', by.y='Mask.ID') # get QC scores
df = merge(struct, imaging, by='MRN')
dim(df)

pairwise.t.test(df$HI_vol_rh.x, df$OLS_HI_categ, p.adjust.method='none')

fm = as.formula(df$HI_vol_rh.x ~ df$OLS_HI_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('volume RH HI, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

Like before, let's clean it up and run it again:

```{r}
bad_nvs = df$OLS_HI_categ == 'NV' & df$OLS_HI_slope.x != 0
df_clean = df[!bad_nvs, ]
library(MASS)
res.lm <- lm(HI_vol_rh.x ~ Sex...Subjects + int_avg_freesurfer5.3 + age_at_scan + I(age_at_scan^2) + ext_avg_freesurfer5.3 + int_avg_freesurfer5.3 + mprage_QC, data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_HI_categ)))
pairwise.t.test(mycluster, df_clean$OLS_HI_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_HI_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('volume RH HI, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

And finally, the logistic regression:

```{r}
library(nnet)
fit <- multinom(df_clean$OLS_HI_categ ~ scale(mycluster), na.action=na.omit)
z <- summary(fit)$coefficients/summary(fit)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
rr = exp(coef(fit))
pp = fitted(fit)
print(p)
print(fit)
print(rr)
```

Now both deterioration and marked improvement are significant when compared to NVs:

* A one-unit increase in the volume of the right hemisphere cluster variable (i.e. increase by 1 SD) is associated with an decrease in the log odds of markedly improving (vs normals) in the amount of .52. Conversely, a one-unit increase is associated with an increase in the log odds of deterioration of .41.
* In terms of relative risk ratio, a one-unit increase in the the volume of that brain cluster yields a relative risk ratio of 1.5 of deterioration, and .6 marked improvement.

## DTI

Let's move on to the DTI clusters:

```{r}
load('~/data/baseline_prediction/combined_descriptives_12172018.RData.gz')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
df = merge(clin, data, by='MRN')
idx = df$diag_group != 'new_onset'
dti = df[!is.na(df$inatt_AD_clu1) & idx,]
load('~/data/baseline_prediction/dti_rd_voxelwise_n272_09212018.RData.gz')
dti = merge(dti, data, by='MRN') # put mask ids in combined dataset
mprage = read.xls('~/data/baseline_prediction/long_scans_08072018.xlsx',
                  sheet='dti')
dti = merge(dti, mprage, by.x='mask.id', by.y='Mask.ID') # get demographics
qc = read.csv('~/data/baseline_prediction/master_qc.csv')
dti = merge(dti, qc, by.x='mask.id', by.y='Mask.ID') # get QC scores
df = merge(dti, imaging, by='MRN')
df$mvmt = rowMeans(scale(df$norm.trans), scale(df$norm.rot))
dim(df)

pairwise.t.test(df$inatt_AD_clu1.x, df$OLS_inatt_categ, p.adjust.method='none')

fm = as.formula(df$inatt_AD_clu1.x ~ df$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('AD inatt clu1, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

Groups are barely different. What if we residualize them?

```{r}
bad_nvs = df$OLS_inatt_categ == 'NV' & df$OLS_inatt_slope.x != 0
df_clean = df[!bad_nvs, ]
library(MASS)
res.lm <- lm(inatt_AD_clu1.x ~ Sex + age_at_scan + I(age_at_scan^2) + mvmt + I(mvmt^2), data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_inatt_categ)))
pairwise.t.test(mycluster, df_clean$OLS_inatt_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('AD inatt clu1, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

Well, group differences did increase when running the brain residuals. Let's check how the logistic model runs:

```{r}
library(nnet)
fit <- multinom(df_clean$OLS_inatt_categ ~ scale(mycluster), na.action=na.omit)
z <- summary(fit)$coefficients/summary(fit)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
rr = exp(coef(fit))
pp = fitted(fit)
print(p)
print(fit)
print(rr)
```

Only the deterioration group is significantly different than NVs:

* A one-unit increase in this first AD cluster is associated with a decrease in the log odds of deterioration in the amount of .49.
* In terms of relative risk ratio, a one-unit decrease in that cluster ields a relative risk ratio of .67 deterioration.

Apparently I didn't even plot the second AD inatt cluster in adhd_NOS_descriptives.md. So, let's switch right to the RD HI cluster:

```{r}
res.lm <- lm(HI_RD_clu1.x ~ Sex + age_at_scan + I(age_at_scan^2) + mvmt + I(mvmt^2), data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_HI_categ)))
pairwise.t.test(mycluster, df_clean$OLS_HI_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_HI_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('RD HI clu1, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

This one looks a bit funky. The only significant difference is between marked and mild??? I'll skip the logistic model for this one.

## rsFMRI

```{r}
load('~/data/baseline_prediction/combined_descriptives_12172018.RData.gz')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
df = merge(clin, data, by='MRN')
idx = df$diag_group != 'new_onset'
fmri = df[!is.na(df$inatt_melodic_limbic) & idx,]
load('~/data/baseline_prediction/melodic_inter_IC11_12142018.RData.gz')
fmri = merge(fmri, data, by='MRN') # put mask ids in combined dataset
mprage = read.xls('~/data/baseline_prediction/long_scans_08072018.xlsx',
                  sheet='mprage')
fmri = merge(fmri, mprage, by.x='mask.id', by.y='Mask.ID...Scan') # get demographics
qc = read.csv('~/data/baseline_prediction/master_qc.csv')
fmri = merge(fmri, qc, by.x='mask.id', by.y='Mask.ID') # get QC scores
df = merge(fmri, imaging, by='MRN')
dim(df)

bad_nvs = df$OLS_inatt_categ == 'NV' & df$OLS_inatt_slope.x != 0
df_clean = df[!bad_nvs, ]
library(MASS)
res.lm <- lm(inatt_melodic_limbic.x ~ age_at_scan + I(age_at_scan^2) + enormGoodTRs_fmri01 + I(enormGoodTRs_fmri01^2) + Sex...Subjects, data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_inatt_categ)))
pairwise.t.test(mycluster, df_clean$OLS_inatt_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('limbic inatt clu, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

```{r}
library(nnet)
fit <- multinom(df_clean$OLS_inatt_categ ~ scale(mycluster), na.action=na.omit)
z <- summary(fit)$coefficients/summary(fit)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
rr = exp(coef(fit))
pp = fitted(fit)
print(p)
print(fit)
print(rr)
```

```{r}
res.lm <- lm(inatt_melodic_VAN.x ~ age_at_scan + I(age_at_scan^2) + enormGoodTRs_fmri01 + I(enormGoodTRs_fmri01^2) + Sex...Subjects, data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_inatt_categ)))
pairwise.t.test(mycluster, df_clean$OLS_inatt_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('VAN inatt clu, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

```{r}
library(nnet)
fit <- multinom(df_clean$OLS_inatt_categ ~ scale(mycluster), na.action=na.omit)
z <- summary(fit)$coefficients/summary(fit)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
rr = exp(coef(fit))
pp = fitted(fit)
print(p)
print(fit)
print(rr)
```

```{r}
res.lm <- lm(inatt_melodic_DMN.x ~ age_at_scan + I(age_at_scan^2) + enormGoodTRs_fmri01 + I(enormGoodTRs_fmri01^2) + Sex...Subjects, data = df_clean)
step <- stepAIC(res.lm, direction = "both", trace = F)
print(summary(step))
mycluster = residuals(step)
summary(aov(lm(mycluster ~ df_clean$OLS_inatt_categ)))
pairwise.t.test(mycluster, df_clean$OLS_inatt_categ, p.adjust.method='none')
fm = as.formula(mycluster ~ df_clean$OLS_inatt_categ)
fit = aov(lm(fm))
boxplot(fm)
title(sprintf('DMN inatt clu, p<%.4f', summary(fit)[[1]][1, 'Pr(>F)']))
```

```{r}
library(nnet)
fit <- multinom(df_clean$OLS_inatt_categ ~ scale(mycluster), na.action=na.omit)
z <- summary(fit)$coefficients/summary(fit)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
rr = exp(coef(fit))
pp = fitted(fit)
print(p)
print(fit)
print(rr)
```


TODO:

* where is that structural cluster actually? (whereami)
* what if we residualize the brain to begin with, using our step function? use lme? 