---
title: "Ordered logistic regression"
output: html_notebook
---

The idea is to run ordered logistic regression, first trying it univariate, then a full model, then the optimized model. Also, do it within domain, and then only the ones that have neuroimaging. Let's start with PRS, which is one of our main targets here.

For the ordinal logistic regression, try the 4 groups, but NV position can be 1 or 4... look at the ANOVA. Philip used 1. Then, try the 3 groups by themselves.

```{r}
prs = read.csv('~/data/PRS_master.csv')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
clin$OLS_inatt_categ = NULL
clin[clin$OLS_inatt_slope <= -.33, 'OLS_inatt_categ'] = 'marked'
clin[clin$OLS_inatt_slope > -.33 & clin$OLS_inatt_slope <= 0, 'OLS_inatt_categ'] = 'mild'
clin[clin$OLS_inatt_slope > 0, 'OLS_inatt_categ'] = 'deter'
clin[clin$DX == 'NV', 'OLS_inatt_categ'] = 'NV'
clin$OLS_inatt_categ = as.factor(clin$OLS_inatt_categ)
clin$OLS_inatt_categ = relevel(clin$OLS_inatt_categ, ref='NV')
clin$OLS_HI_categ = NULL
clin[clin$OLS_HI_slope <= -.5, 'OLS_HI_categ'] = 'marked'
clin[clin$OLS_HI_slope > -.5 & clin$OLS_HI_slope <= 0, 'OLS_HI_categ'] = 'mild'
clin[clin$OLS_HI_slope > 0, 'OLS_HI_categ'] = 'deter'
clin[clin$DX == 'NV', 'OLS_HI_categ'] = 'NV'
clin$OLS_HI_categ = as.factor(clin$OLS_HI_categ)
clin$OLS_HI_categ = relevel(clin$OLS_HI_categ, ref='NV')

clin$HI_ordered = factor(clin$OLS_HI_categ, levels=c('NV', 'marked', 'mild', 'deter'), ordered=T)
clin$inatt_ordered = factor(clin$OLS_inatt_categ, levels=c('NV', 'marked', 'mild', 'deter'), ordered=T)

df = merge(prs, clin, by='MRN', all.x=F, all.y=F)
```

Now, let's check each PRS individually:

```{r}
library(MASS)

vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=sum(vars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in colnames(prs)[vars]) {
    fm_str = sprintf('%s ~ scale(%s)', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = colnames(prs)[vars]
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRS_univariate_everyone.csv', target)
  write.csv(res, file=out_fname)
}
```

But we can also chuck in a model that has everything in it:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRS_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

And finally a stepwise best model approach:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRS_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Now, one could argue that we're shooting ourselves in the foot by throwing crappy PRS (ASD, SCZ) in the model. So, let's re-run the two big models with the ADHD PRS only.

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRSADHD_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRSADHD_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# neuroimaging only

Now we repeat all the steps above, but only looking at the kids that had neuroimaging:

```{r}
prs = read.csv('~/data/PRS_master.csv')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
clin$OLS_inatt_categ = NULL
clin[clin$OLS_inatt_slope <= -.33, 'OLS_inatt_categ'] = 'marked'
clin[clin$OLS_inatt_slope > -.33 & clin$OLS_inatt_slope <= 0, 'OLS_inatt_categ'] = 'mild'
clin[clin$OLS_inatt_slope > 0, 'OLS_inatt_categ'] = 'deter'
clin[clin$DX == 'NV', 'OLS_inatt_categ'] = 'NV'
clin$OLS_inatt_categ = as.factor(clin$OLS_inatt_categ)
clin$OLS_inatt_categ = relevel(clin$OLS_inatt_categ, ref='NV')
clin$OLS_HI_categ = NULL
clin[clin$OLS_HI_slope <= -.5, 'OLS_HI_categ'] = 'marked'
clin[clin$OLS_HI_slope > -.5 & clin$OLS_HI_slope <= 0, 'OLS_HI_categ'] = 'mild'
clin[clin$OLS_HI_slope > 0, 'OLS_HI_categ'] = 'deter'
clin[clin$DX == 'NV', 'OLS_HI_categ'] = 'NV'
clin$OLS_HI_categ = as.factor(clin$OLS_HI_categ)
clin$OLS_HI_categ = relevel(clin$OLS_HI_categ, ref='NV')

clin$HI_ordered = factor(clin$OLS_HI_categ, levels=c('NV', 'marked', 'mild', 'deter'), ordered=T)
clin$inatt_ordered = factor(clin$OLS_inatt_categ, levels=c('NV', 'marked', 'mild', 'deter'), ordered=T)

df = merge(prs, clin, by='MRN', all.x=F, all.y=F)

img = read.csv('~/data/baseline_prediction/residualized_combined_OLS_descriptives_02072019.csv')
df = merge(df, img, by='MRN', all.x=F, all.y=F)
```

Like before, run univariate PRS:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=sum(vars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in colnames(prs)[vars]) {
    fm_str = sprintf('%s ~ scale(%s)', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = colnames(prs)[vars]
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRS_univariate_imaging.csv', target)
  write.csv(res, file=out_fname)
}
```

Chuck in a model that has everything in it:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRS_full_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Stepwise best model approach:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRS_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

And we repeat it with the ADHD PRS only.

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRSADHD_full_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_PRSADHD_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Brain

Let's do a similar approach and run the brain univariate, full, and stepwise. For full and stepwise, let's play with the intersection of all neuroimaging for now. If that doesn't work, or the sample is too small, we can impute it.

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
myvars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=length(myvars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in myvars) {
    fm_str = sprintf('%s ~ scale(%s)', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = myvars
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_imaging_univariate_everyone.csv', target)
  write.csv(res, file=out_fname)
}
```

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
keep_me = rowSums(is.na(df))==0
df_inter = df[keep_me, ]

myvars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df_inter[, myvars]))
  mydata[, target] = df_inter[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:length(myvars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_imaging_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Finally, we run stepwise:

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
keep_me = rowSums(is.na(df))==0
df_inter = df[keep_me, ]

myvars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df_inter[, myvars]))
  mydata[, target] = df_inter[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 0) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_imaging_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Sex, age, IQ, and baseline SX

We should have those variables for everyone in the cohort. Again, let's do the univariate approach, then full model, followed by stepwise.

```{r}
load('~/data/baseline_prediction/adhd200_10042018.RData.gz')
df = merge(data, clin, by='MRN', all.x=F, all.y=F)

# remove handedness because of NAs
df$vCateg_handedness = NULL

myvars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=length(myvars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in myvars) {
    fm_str = sprintf('%s ~ %s', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = myvars
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_others_univariate_everyone.csv', target)
  write.csv(res, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

myvars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(df_clean[, myvars])
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:length(myvars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_others_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

myvars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(df_clean[, myvars])
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 0) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_others_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Combining everything

The best approach here is to start with imaging intersection, and attach everything to that. Then, we run the univariate (just to make sure past results still somewhat hold), full model and stepwise.

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
keep_me = rowSums(is.na(df))==0
df_inter = df[keep_me, ]

df = merge(df_inter, prs, by='MRN', all.x=F, all.y=F)
df = merge(df, data, by='MRN', all.x=F, all.y=F)
# remove handedness because of NAs
df$vCateg_handedness = NULL

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.01.profile', 'PROFILES.0.0001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.05.profile', 'PROFILES.0.00000005.profile', 'PROFILES.0.01.profile',
                 'PROFILES.0.0000001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }

  res = matrix(nrow=length(myvars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in myvars) {
    fm_str = sprintf('%s ~ %s', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = myvars
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_all_univariate_imaging.csv', target)
  write.csv(res, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.01.profile', 'PROFILES.0.0001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.05.profile', 'PROFILES.0.00000005.profile', 'PROFILES.0.01.profile',
                 'PROFILES.0.0000001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  mydata = as.data.frame(scale(df_clean[, c(brain_vars, prs_vars)]))
  mydata = cbind(mydata, as.data.frame(df_clean[, other_vars]))
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:length(myvars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_all_full_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.01.profile', 'PROFILES.0.0001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.05.profile', 'PROFILES.0.00000005.profile', 'PROFILES.0.01.profile',
                 'PROFILES.0.0000001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  mydata = as.data.frame(scale(df_clean[, c(brain_vars, prs_vars)]))
  mydata = cbind(mydata, as.data.frame(df_clean[, other_vars]))
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 1) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_all_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Out of curiosity, is we don't include the other two rsFMRI patterns, does DMN get selected?

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

brain_vars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_DMN")

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.01.profile', 'PROFILES.0.0001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.05.profile', 'PROFILES.0.00000005.profile', 'PROFILES.0.01.profile',
                 'PROFILES.0.0000001.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  mydata = as.data.frame(scale(df_clean[, c(brain_vars, prs_vars)]))
  mydata = cbind(mydata, as.data.frame(df_clean[, other_vars]))
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 1) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_allDMNonly_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Clinical groups only

What if we repeat this entire analysis with the clinical group only? This way the brain wouldn't be penalized for having NVs smack in the middle. Let's see how it goes:


```{r}
prs = read.csv('~/data/PRS_master.csv')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
clin = clin[clin$DX != 'NV', ]
clin$OLS_inatt_categ = NULL
clin[clin$OLS_inatt_slope <= -.33, 'OLS_inatt_categ'] = 'marked'
clin[clin$OLS_inatt_slope > -.33 & clin$OLS_inatt_slope <= 0, 'OLS_inatt_categ'] = 'mild'
clin[clin$OLS_inatt_slope > 0, 'OLS_inatt_categ'] = 'deter'
clin$OLS_inatt_categ = as.factor(clin$OLS_inatt_categ)
clin$OLS_HI_categ = NULL
clin[clin$OLS_HI_slope <= -.5, 'OLS_HI_categ'] = 'marked'
clin[clin$OLS_HI_slope > -.5 & clin$OLS_HI_slope <= 0, 'OLS_HI_categ'] = 'mild'
clin[clin$OLS_HI_slope > 0, 'OLS_HI_categ'] = 'deter'
clin$OLS_HI_categ = as.factor(clin$OLS_HI_categ)

clin$HI_ordered = factor(clin$OLS_HI_categ, levels=c('marked', 'mild', 'deter'), ordered=T)
clin$inatt_ordered = factor(clin$OLS_inatt_categ, levels=c('marked', 'mild', 'deter'), ordered=T)

df = merge(prs, clin, by='MRN', all.x=F, all.y=F)
```

Now, let's check each PRS individually:

```{r}
library(MASS)

vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=sum(vars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in colnames(prs)[vars]) {
    fm_str = sprintf('%s ~ scale(%s)', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = colnames(prs)[vars]
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRS_univariate_everyone.csv', target)
  write.csv(res, file=out_fname)
}
```

But we can also chuck in a model that has everything in it:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRS_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

And finally a stepwise best model approach:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRS_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Now, one could argue that we're shooting ourselves in the foot by throwing crappy PRS (ASD, SCZ) in the model. So, let's re-run the two big models with the ADHD PRS only.

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRSADHD_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRSADHD_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# neuroimaging only

Now we repeat all the steps above, but only looking at the kids that had neuroimaging:

```{r}
prs = read.csv('~/data/PRS_master.csv')
clin = read.csv('~/data/baseline_prediction/long_clin_11302018.csv')
clin = clin[clin$DX != 'NV', ]
clin$OLS_inatt_categ = NULL
clin[clin$OLS_inatt_slope <= -.33, 'OLS_inatt_categ'] = 'marked'
clin[clin$OLS_inatt_slope > -.33 & clin$OLS_inatt_slope <= 0, 'OLS_inatt_categ'] = 'mild'
clin[clin$OLS_inatt_slope > 0, 'OLS_inatt_categ'] = 'deter'
clin$OLS_inatt_categ = as.factor(clin$OLS_inatt_categ)
clin$OLS_HI_categ = NULL
clin[clin$OLS_HI_slope <= -.5, 'OLS_HI_categ'] = 'marked'
clin[clin$OLS_HI_slope > -.5 & clin$OLS_HI_slope <= 0, 'OLS_HI_categ'] = 'mild'
clin[clin$OLS_HI_slope > 0, 'OLS_HI_categ'] = 'deter'
clin$OLS_HI_categ = as.factor(clin$OLS_HI_categ)

clin$HI_ordered = factor(clin$OLS_HI_categ, levels=c('marked', 'mild', 'deter'), ordered=T)
clin$inatt_ordered = factor(clin$OLS_inatt_categ, levels=c('marked', 'mild', 'deter'), ordered=T)

df = merge(prs, clin, by='MRN', all.x=F, all.y=F)

img = read.csv('~/data/baseline_prediction/residualized_combined_OLS_descriptives_02072019.csv')
df = merge(df, img, by='MRN', all.x=F, all.y=F)
```

Like before, run univariate PRS:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=sum(vars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in colnames(prs)[vars]) {
    fm_str = sprintf('%s ~ scale(%s)', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = colnames(prs)[vars]
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRS_univariate_imaging.csv', target)
  write.csv(res, file=out_fname)
}
```

Chuck in a model that has everything in it:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRS_full_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Stepwise best model approach:

```{r}
vars = grepl(pattern='^PROFI', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRS_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

And we repeat it with the ADHD PRS only.

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:sum(vars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRSADHD_full_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
vars = grepl(pattern='file$', colnames(prs))
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df[, colnames(prs)[vars]]))
  mydata[, target] = df[, target]
  fm_str = sprintf('%s ~ .', target, v)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  ctable <- coef(summary(myfit))[1:nvars, ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_PRSADHD_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Brain

Let's do a similar approach and run the brain univariate, full, and stepwise. For full and stepwise, let's play with the intersection of all neuroimaging for now. If that doesn't work, or the sample is too small, we can impute it.

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
myvars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=length(myvars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in myvars) {
    fm_str = sprintf('%s ~ scale(%s)', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = myvars
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_imaging_univariate_everyone.csv', target)
  write.csv(res, file=out_fname)
}
```

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
keep_me = rowSums(is.na(df))==0
df_inter = df[keep_me, ]

myvars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df_inter[, myvars]))
  mydata[, target] = df_inter[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:length(myvars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_imaging_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Finally, we run stepwise:

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
keep_me = rowSums(is.na(df))==0
df_inter = df[keep_me, ]

myvars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(scale(df_inter[, myvars]))
  mydata[, target] = df_inter[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 0) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_imaging_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Sex, age, IQ, and baseline SX

We should have those variables for everyone in the cohort. Again, let's do the univariate approach, then full model, followed by stepwise.

```{r}
load('~/data/baseline_prediction/adhd200_10042018.RData.gz')
df = merge(data, clin, by='MRN', all.x=F, all.y=F)

# remove handedness because of NAs
df$vCateg_handedness = NULL

myvars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
for (target in c('inatt_ordered', 'HI_ordered')) {
  res = matrix(nrow=length(myvars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in myvars) {
    fm_str = sprintf('%s ~ %s', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = myvars
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_others_univariate_everyone.csv', target)
  write.csv(res, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

myvars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(df_clean[, myvars])
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:length(myvars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_others_full_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

myvars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
for (target in c('inatt_ordered', 'HI_ordered')) {
  mydata = as.data.frame(df_clean[, myvars])
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 0) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_others_step_everyone.csv', target)
  write.csv(ctable, file=out_fname)
}
```

# Combining everything

The best approach here is to start with imaging intersection, and attach everything to that. Then, we run the univariate (just to make sure past results still somewhat hold), full model and stepwise.

```{r}
df = merge(img, clin, by='MRN', all.x=F, all.y=F)
keep_me = rowSums(is.na(df))==0
df_inter = df[keep_me, ]

df = merge(df_inter, prs, by='MRN', all.x=F, all.y=F)
df = merge(df, data, by='MRN', all.x=F, all.y=F)

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile', 'PROFILES.0.2.profile', 'PROFILES.0.4.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  res = matrix(nrow=length(myvars), ncol=8)
  colnames(res) = c('coef', 'tstat', 'pval', 'coef2.5CI', 'coef97.5CI', 'OR', 'OR2.5CI', 'OR97.5CI')
  cnt = 1
  for (v in myvars) {
    fm_str = sprintf('%s ~ %s', target, v) 
    myfit <- polr(as.formula(fm_str), data=df, Hess=TRUE)
    ctable <- coef(summary(myfit))
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    res[cnt, 'coef'] = ctable[1, 'Value']
    res[cnt, 'tstat'] = ctable[1, 't value']
    res[cnt, 'pval'] = p[1]
    
    ci<-confint.default(myfit)
    res[cnt, c('coef2.5CI', 'coef97.5CI')] = ci
    ors = exp(cbind(OR = coef(myfit), ci))
    res[cnt, c('OR', 'OR2.5CI', 'OR97.5CI')] = ors
    cnt = cnt + 1
  }
  rownames(res) = myvars
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_all_univariate_imaging.csv', target)
  write.csv(res, file=out_fname)
}
```

Here we'll be down to 80 subjects (from 103 that have all 3 modalities, 101 that we ran for univariate). The full model doesn't run here. So, let's remove some of the variables that didnt work so well in the univariate results:

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile', 'PROFILES.0.2.profile', 'PROFILES.0.4.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  mydata = as.data.frame(scale(df_clean[, c(brain_vars, prs_vars)]))
  mydata = cbind(mydata, as.data.frame(df_clean[, other_vars]))
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  ctable <- coef(summary(myfit))[1:length(myvars), ]
  p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
  ctable <- cbind(ctable, "p value" = p)
  ci<-confint.default(myfit)
  ctable <- cbind(ctable, ci)
  ors = exp(cbind(OR = coef(myfit), ci))
  colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
  ctable <- cbind(ctable, ors)
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_all_full_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_limbic", "inatt_VAN", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile', 'PROFILES.0.2.profile', 'PROFILES.0.4.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  mydata = as.data.frame(scale(df_clean[, c(brain_vars, prs_vars)]))
  mydata = cbind(mydata, as.data.frame(df_clean[, other_vars]))
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 0) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_all_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```

Out of curiosity, is we don't include the other two rsFMRI patterns, does DMN get selected?

```{r}
keep_me = rowSums(is.na(df))==0
df_clean = df[keep_me, ]

brain_vars = c("inatt_struct", "HI_struct", "inatt_AD", "inatt_DMN")

for (target in c('inatt_ordered', 'HI_ordered')) {
  myvars = c()
  if (target == 'inatt_ordered') {
    brain_vars = c("inatt_struct", "inatt_AD", "inatt_DMN")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile', 'PROFILES.0.2.profile', 'PROFILES.0.4.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  } else {
    brain_vars = c("HI_struct")
    for (s in brain_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    prs_vars = c('PROFILES.0.5.profile')
    for (s in prs_vars) {
      myvars = c(myvars, sprintf('scale(%s)', s))
    }
    other_vars = c("vCateg_sex", "v_Age", "v_IQ", "SX_inatt_baseline", "SX_HI_baseline")
    myvars = c(myvars, other_vars)
  }
  mydata = as.data.frame(scale(df_clean[, c(brain_vars, prs_vars)]))
  mydata = cbind(mydata, as.data.frame(df_clean[, other_vars]))
  mydata[, target] = df_clean[, target]
  fm_str = sprintf('%s ~ .', target)
  myfit <- polr(as.formula(fm_str), data=mydata, Hess=TRUE)
  step <- stepAIC(myfit, direction = "both", trace = F)
  myfit = step
  nvars = length(coef(myfit))
  if (nvars > 0) { 
    ctable <- coef(summary(myfit))[1:nvars, ]
    p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
    ctable <- cbind(ctable, "p value" = p)
    ci<-confint.default(myfit)
    ctable <- cbind(ctable, ci)
    ors = exp(cbind(OR = coef(myfit), ci))
    colnames(ors)[2:3] = c('OR2.5CI', 'OR97.5CI')
    ctable <- cbind(ctable, ors)
  } else {
    ctable = 0
  }
    
  out_fname = sprintf('~/data/baseline_prediction/logistic_ordinal_results/%s_clinicalOnly_allDMNonly_step_imaging.csv', target)
  write.csv(ctable, file=out_fname)
}
```


# TODO
* Complete IQ variable

