---
title: "Removing correlation for baseline prediction study after fixing transformation variables"
output:
  html_notebook: default
  html_document: default
---

# 09/18/2018

We're checking this again since we changed the transformation variable we read from TORTOISE. So, this is a continuation of the _principled note.

```{r}
library(gdata)
demo = read.xls('~/data/baseline_prediction/long_scans_08072018.xlsx', sheet='dti')
df = read.csv('~/data/baseline_prediction/dti_long_09182018.csv')
colnames(demo)[1] ='MRN'
df2 = merge(df, demo, by.x = 'mask.id', by.y='Mask.ID')
dim(df2)
```

Now remove outliers in each of the normal distributions:

```{r}
mult = 2
hist(df2$mean_fa, breaks=50)
ll = mean(df2$mean_fa) - mult*sd(df2$mean_fa)
ul = mean(df2$mean_fa) + mult*sd(df2$mean_fa)
abline(v = ll, col='red')
abline(v = ul, col='red')
idx_fa = df2$mean_fa < ul & df2$mean_fa > ll
```
```{r}
mult = 2
hist(df2$mean_ad, breaks=50)
ll = mean(df2$mean_ad) - mult*sd(df2$mean_ad)
ul = mean(df2$mean_ad) + mult*sd(df2$mean_ad)
abline(v = ll, col='red')
abline(v = ul, col='red')
idx_ad = df2$mean_ad < ul & df2$mean_ad > ll
```

```{r}
mult = 2
hist(df2$mean_rd, breaks=50)
ll = mean(df2$mean_rd) - mult*sd(df2$mean_rd)
ul = mean(df2$mean_rd) + mult*sd(df2$mean_rd)
abline(v = ll, col='red')
abline(v = ul, col='red')
idx_rd = df2$mean_rd < ul & df2$mean_rd > ll
```

```{r}
idx = idx_fa & idx_ad & idx_rd
df3 = df2[idx,]
dim(df3)
```

Maybe we also include a check on goodVolumes?

```{r}
mult = 2
hist(df3$goodVolumes, breaks=50)
abline(v = 58, col='red')
df3 = df3[df3$goodVolumes >=58,]
```

OK, now we have removed the bad scans based on distribution. Now, let's pick the first one for each subject, and remove them sequentially according to movement:

```{r}
source('~/research_code/baseline_prediction/aux_functions.R', echo=TRUE)
df = get_single_scans(df3, last_scan=F)
dim(df)
cor.test(df$norm.trans, df$mean_fa)
cor.test(df$norm.trans, df$mean_ad)
cor.test(df$norm.trans, df$mean_rd)
cor.test(df$norm.rot, df$mean_fa)
cor.test(df$norm.rot, df$mean_ad)
cor.test(df$norm.rot, df$mean_rd)
```

Actually, there is some correlation here. Before we remove stuff, let's merge it with clinical because maybe after removing other scans some things might change:

```{r}
dim(df)
clin = read.csv('~/data/baseline_prediction//long_clin_0918.csv')
dim(clin)
m = merge(df,clin)
dim(m)
scan_date = as.Date(as.character(m$record.date.collected...Scan), format='%m/%d/%Y')
doa_date = as.Date(as.character(m$last_DOA), format='%m/%d/%y')
d = doa_date - scan_date
sum(d > 30*6)
sum(d > 30*12)
```

OK, so let's see if we have correlations using any of those subject sets:

```{r}
cor.test(m$norm.trans, m$mean_fa)
cor.test(m$norm.trans, m$mean_ad)
cor.test(m$norm.trans, m$mean_rd)
cor.test(m$norm.rot, m$mean_fa)
cor.test(m$norm.rot, m$mean_ad)
cor.test(m$norm.rot, m$mean_rd)
```

Yep, still some correlation. Let's look specifically at only the scans with at least one year difference.

```{r}
idx = d > 30*12
sum(idx)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_rd)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_rd)
```

OK, so what can we remove to reduce this correlation?

```{r}
hist(m[idx,]$norm.trans, breaks=50)
```

We are going to remove too much if we do it here. Let's go back to the original data:

```{r}
mult = 2
hist(df3$norm.rot, breaks=50)
ll = mean(df3$norm.rot) - mult*sd(df3$norm.rot)
ul = mean(df3$norm.rot) + mult*sd(df3$norm.rot)
abline(v = ll, col='red')
abline(v = ul, col='red')
idx_rot = df3$norm.rot < ul & df3$norm.rot > ll
```

```{r}
mult = 2
hist(df3$norm.trans, breaks=50)
ll = mean(df3$norm.trans) - mult*sd(df3$norm.trans)
ul = mean(df3$norm.trans) + mult*sd(df3$norm.trans)
abline(v = ll, col='red')
abline(v = ul, col='red')
idx_trans = df3$norm.trans < ul & df3$norm.trans > ll
```

I could actually even be a bit more conservative for these thresholds if I go by eye, but let's see what we get just by using the principled way:

```{r}
df = get_single_scans(df3[idx_rot & idx_trans, ], last_scan=F)
m = merge(df,clin)
scan_date = as.Date(as.character(m$record.date.collected...Scan), format='%m/%d/%Y')
doa_date = as.Date(as.character(m$last_DOA), format='%m/%d/%y')
d = doa_date - scan_date
idx = d > 30*12
sum(idx)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_rd)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_rd)
```

Still no good. Let's be more stringent then:

```{r}
# hist(df3$norm.rot, breaks=50)
ul = .016
# abline(v = ul, col='red')
idx_rot = df3$norm.rot < ul

# hist(df3$norm.trans, breaks=50)
ul = 1.5
# abline(v = ul, col='red')
idx_trans = df3$norm.trans < ul

df = get_single_scans(df3[idx_rot & idx_trans, ], last_scan=F)
m = merge(df,clin)
scan_date = as.Date(as.character(m$record.date.collected...Scan), format='%m/%d/%Y')
doa_date = as.Date(as.character(m$last_DOA), format='%m/%d/%y')
d = doa_date - scan_date
idx = d > 30*12
sum(idx)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_rd)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_rd)
```

So, this worked, but now I've slashed my data from 263 to 180 subjects... let's see if we go back to a less-principled way to find thresholds, and increase these numbers:

```{r}
idx_fa = df2$mean_fa < .5 & df2$mean_fa > .42
idx_ad = df2$mean_ad < 1.4 & df2$mean_ad > 1.2
idx_rd = df2$mean_rd < .62 & df2$mean_rd > .52
idx_rot = df2$norm.rot < .04
idx_trans = df2$norm.trans < 2
idx_gv = df2$goodVolumes >= 58

df = get_single_scans(df2[idx_rot & idx_trans & idx_fa & idx_ad & idx_rd & idx_gv, ], last_scan=F)
m = merge(df,clin)
scan_date = as.Date(as.character(m$record.date.collected...Scan), format='%m/%d/%Y')
doa_date = as.Date(as.character(m$last_DOA), format='%m/%d/%y')
d = doa_date - scan_date
idx = d > 30*12
sum(idx)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_rd)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_rd)
```

These are better... 223 subjects is a bit better. Now, need to re-run the dti results.

Now, say I only remove based on principled thresholds. How does it look? First, the movement variables are not very normal, and also I don't see it as bad having little movement. So, let's make a few modification there:

```{r}
hist(log(df3$norm.rot), breaks=50)
hist(log(df3$norm.trans), breaks=50)
```

```{r}
hist(scale(log(df3$norm.rot)), breaks=50)
```

```{r}
mult = 2
idx_fa = df2$mean_fa < (mean(df2$mean_fa) + mult*sd(df2$mean_fa)) & (df2$mean_fa > mean(df2$mean_fa) - mult*sd(df2$mean_fa))
idx_ad = df2$mean_ad < (mean(df2$mean_ad) + mult*sd(df2$mean_ad)) & df2$mean_ad > (mean(df2$mean_ad) - mult*sd(df2$mean_ad))
idx_rd = df2$mean_rd < (mean(df2$mean_rd) + mult*sd(df2$mean_rd)) & df2$mean_rd > mean(df2$mean_rd) - mult*sd(df2$mean_rd)
idx_rot = log(df2$norm.rot) < (mean(log(df2$norm.rot) + mult*sd(log(df2$norm.rot))))
idx_trans = log(df2$norm.trans) < (mean(log(df2$norm.trans) + mult*sd(log(df2$norm.trans))))

df = get_single_scans(df2[idx_rot & idx_trans & idx_fa & idx_ad & idx_rd, ], last_scan=F)
m = merge(df,clin)
scan_date = as.Date(as.character(m$record.date.collected...Scan), format='%m/%d/%Y')
doa_date = as.Date(as.character(m$last_DOA), format='%m/%d/%y')
d = doa_date - scan_date
idx = d > 30*12
sum(idx)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.trans, m[idx,]$mean_rd)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_fa)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_ad)
cor.test(m[idx,]$norm.rot, m[idx,]$mean_rd)
```

So, as expected we still have lots of correlation. However, can we check each voxel, and just remove those voxels that have some correlation to the variable? This way we could potentially still keep our 272 subjects. Philip didn't seem crazy about this idea, but it's worth a try. We could actually do it both way (272 and 223) and then just check the correlation of the good features with movement.

We should also make sure there is no correlation between movement and the variables of interest!

```{r}
m272 = m[idx,]
dim(m272)

idx_fa = df2$mean_fa < .5 & df2$mean_fa > .42
idx_ad = df2$mean_ad < 1.4 & df2$mean_ad > 1.2
idx_rd = df2$mean_rd < .62 & df2$mean_rd > .52
idx_rot = df2$norm.rot < .04
idx_trans = df2$norm.trans < 2
idx_gv = df2$goodVolumes >= 58

df = get_single_scans(df2[idx_rot & idx_trans & idx_fa & idx_ad & idx_rd & idx_gv, ], last_scan=F)
m = merge(df,clin)
scan_date = as.Date(as.character(m$record.date.collected...Scan), format='%m/%d/%Y')
doa_date = as.Date(as.character(m$last_DOA), format='%m/%d/%y')
d = doa_date - scan_date
idx = d > 30*12
m223 = m[idx, ]
dim(m223)
```

```{r}
cor.test(m272$norm.trans, m272$OLS_inatt_slope)
cor.test(m272$norm.trans, m272$OLS_HI_slope)
cor.test(m272$norm.trans, m272$OLS_total_slope)
cor.test(m272$norm.trans, m272$random_inatt_slope)
cor.test(m272$norm.trans, m272$random_hi_slope)
cor.test(m272$norm.trans, m272$random_total_slope)
pairwise.t.test(m272$norm.trans, m272$diag_group2)
pairwise.t.test(m272$norm.trans, m272$group_INATT3)
pairwise.t.test(m272$norm.trans, m272$group_HI3)
pairwise.t.test(m272$norm.trans, m272$group_total3)

cor.test(m272$norm.rot, m272$OLS_inatt_slope)
cor.test(m272$norm.rot, m272$OLS_HI_slope)
cor.test(m272$norm.rot, m272$OLS_total_slope)
cor.test(m272$norm.rot, m272$random_inatt_slope)
cor.test(m272$norm.rot, m272$random_hi_slope)
cor.test(m272$norm.rot, m272$random_total_slope)
pairwise.t.test(m272$norm.rot, m272$diag_group2)
pairwise.t.test(m272$norm.rot, m272$group_INATT3)
pairwise.t.test(m272$norm.rot, m272$group_HI3)
pairwise.t.test(m272$norm.rot, m272$group_total3)
```

```{r}
cor.test(m223$norm.trans, m223$OLS_inatt_slope)
cor.test(m223$norm.trans, m223$OLS_HI_slope)
cor.test(m223$norm.trans, m223$OLS_total_slope)
cor.test(m223$norm.trans, m223$random_inatt_slope)
cor.test(m223$norm.trans, m223$random_hi_slope)
cor.test(m223$norm.trans, m223$random_total_slope)
pairwise.t.test(m223$norm.trans, m223$diag_group2)
pairwise.t.test(m223$norm.trans, m223$group_INATT3)
pairwise.t.test(m223$norm.trans, m223$group_HI3)
pairwise.t.test(m223$norm.trans, m223$group_total3)

cor.test(m223$norm.rot, m223$OLS_inatt_slope)
cor.test(m223$norm.rot, m223$OLS_HI_slope)
cor.test(m223$norm.rot, m223$OLS_total_slope)
cor.test(m223$norm.rot, m223$random_inatt_slope)
cor.test(m223$norm.rot, m223$random_hi_slope)
cor.test(m223$norm.rot, m223$random_total_slope)
pairwise.t.test(m223$norm.rot, m223$diag_group2)
pairwise.t.test(m223$norm.rot, m223$group_INATT3)
pairwise.t.test(m223$norm.rot, m223$group_HI3)
pairwise.t.test(m223$norm.rot, m223$group_total3)
```



