---
title: "Subgroup analysis"
output: html_notebook
---

# 2018-09-19 14:30:54

Let's re-run the analysis for different subgroups. I'll change the code to add
different cases when restricting the groups, but in some cases the different
targets won't apply.

In particular, I don't think it's valid to run the LCMM groups, as we'll be
removing subjects and the LCMM groups were created using all of them. I can see
an aergument for doing it anyways, but for now let's avoid it. The same idea
applies to not using random slopes.

## ADHD

Here we remove anyone who is unaffected (diag_group). That leaves people in full and partial
remission, as well as persistent and new_onset. We can run OLS regression, do a
3 group classification, 2 group (persistent vs remission without new onset).

```{r}
df = read.csv('~/data/baseline_prediction/long_clin_0918.csv')
df = df[df$diag_group != 'unaffect', ]
df$diag_group = factor(df$diag_group)
table(df$diag_group)
table(df$diag_group2)
```

## ADHD_NOS

Here we remove anyone who is NV (DX). That leaves people who are ADHD and
ADHD_NOS. In this case, the labels of remission and persistence don't apply, so
we have to categorize based on OLS slopes. We can then have 2 groups (improvers
vs non-improvers), or do the OLS regression.

```{r}
df = read.csv('~/data/baseline_prediction/long_clin_0918.csv')
df = df[df$DX != 'NV', ]
plot(df$OLS_HI_slope)
```

## No new-onset

Here we remove anyone who had late-onset ADHD (diag_group). We can still run OLS slopes, and
and diag_group2. If we categorize slopes we'd have 2 groups, but NV would be
mixed with non-improvers. It's still worth running it, just to compare with the entire group analysis. Or, we could also disregard NVs and just run improvers vs
non-improvers. 

```{r}
df = read.csv('~/data/baseline_prediction/long_clin_0918.csv')
df = df[df$diag_group != 'new_onset', ]
df$diag_group = factor(df$diag_group)
table(df$diag_group)
table(df$diag_group2)
plot(df$OLS_HI_slope)
```